Preventing a Nuclear ISIL: \\Recommendations for United States Policy

### 2015-05-10 draft.tex
Finished chopping up old Chpater 6 material.

Assuming that the White House institutionalizes this role as a Deputy National Security Advisor, the costs associated with this recommendation equal the salary and benefits of this person as well as a small staff. There will also be costs associated with the initiatives created by this person and other staff time that is devoted to those initiatives. Creating this role is an important step in crafting a strategic plan that integrates departmental efforts. The downside to this recommendation is that the efforts of this person are likely to be hampered by turf wars inherent in integrating departmental efforts; however, this difficulty can be mitigated by an approach that seeks to provide departments with a framework for evaluating and adjusting their efforts in the context of the broader effort, rather than mandating changes from the top down \cite{Levi2008}.

@@

@@

@@

\section{???????  Old Chapter 6 Material follows  ??????}

\subsection{????????  Orphan Policy Statements, need polishing and integration ????????}

This report will concentrate on overcoming the complexities in the IC organizational structure to facilitate information sharing. The policy recommendation is a two-pronged strategy that addresses intelligence sharing and analysis on both the domestic and international aspects. The goal for the intelligence community is to foster efficient coordination by improving the flow of information among intelligence organizations, changing the \enquote{need-to-know} basis culture of information sharing, and promoting integration across the organizations to provide time-pertinent. In terms of the international side, the policy will revolve around identifying countries and international bodies the U.S is willing to share information with and to what extent the parties will be involved in the process of intelligence analysis and policy formation.

\subsection{Intel Policy Solution}

The best means of counteracting this fragmentation is to develop a more integrated system with the ODNI at the helm of this  centralized structure. The Office of the Director of National Intelligence should be allowed more authority to dictate the management of information across agencies.  To increase the coordination and familiarity with other agencies in the intelligence community, personnel should do rotations in each of the agencies. This way, it promotes cross-agency relationships and personnel become better equipped at sharing and comparing methods of collecting and analyzing information.  Agencies should be less structured on rules and regulations and instead focus on how to best adapt to changing circumstances. A combination of minimal oversight and conflicting missions due the independent nature of these agencies \cite{Zegart2005} require that ODNI implement a uniformity of standards to decrease costs and simplify the flow of information.

\subsection{Military Policy Solution}

As ISIL carves its traction in Iraq and Syria, the two countries become an immediate concern for help by the U.S.. Strengthening Iraq and Syria could be among the first line of defense against ISIL. As U.S. has an improving bilateral relations with Iraq, Iraq become an easier partner for U.S. to work with in reigning ISIL. The Strategic Framework Agreement for a Relationship of Friendship and Cooperation between the U.S. and the Republic of Iraq guides the overall political, economic, cultural, and security ties with Iraq.  This agreement is designed to help the Iraqi people stand on their own and reinforce Iraqi sovereignty, while protecting U.S. interests in the Middle East.

However, the country's relation with Syria is not black-and-white as we would like it to be. The U.S.-Syria relationship is negative after having suspended diplomatic relations in 2014. The U.S. has also imposed economic sanction to Syria since 2004 due to the country's involvement in harboring terrorist groups and the meddling of Lebanon's political affairs including the assassination of Lebanese Prime Minister Rafik Hariri. The U.S. deeply opposed to the ongoing Syrian Civil war and President Bashar al-Assad's regime, specifically regarding the country's use and stockpile of chemical and biological weapons. The U.S. has floated the idea to train the Syrian military opposition, and even considering to train the Syrian military itself in order to fight ISIL, nevertheless, President Assad viewed the proposal as infringing on their country's sovereignty. Despite the two countries differences, they actually both share a common enemy, and it is hope this fact can be harnessed to U.S.' advantage.

\subsubsection{????????  Orphan Policy Statements, need polishing and integration ????????}

\subsection{Proliferation of RDD materials}

Interception (proliferation)

\subsubsection{Fission Weapons}

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is widely believed that it has nuclear capabilities. Turkey is a NATO nuclear weapon sharing state and has U.S. tactical nuclear weapons stationed within its borders. However, it is unlikely that ISIL would be able to steal a nuclear weapon from either of these states, as those in Turkey are under U.S. protection and Israel has proven adept at hiding their arsenal. Of greatest concern is the possibility that any of these nuclear-capable states, North Korea in particular, would willingly sell a physics package to ISIL.

The non-proliferation treaty (NPT) assists with prevention of these nations selling or having their weapons stolen. The NPT has a multitude of safeguards in place, including accounting for the flow of materials in and out of nuclear facilities, physical security, and containment and surveillance.However, while the NPT's has been effective in preventing nations from using existing energy facilities to make off the books weapon grade materials (these would be easier to sell), Iran, Iraq, and North Korea were able to set up separate facilities for creating weapons grade material. While these facilities were set up without international knowledge, they were discovered in Iraq and North Korea rather quickly.

\subsubsection{RDD}
 Some of the materials required to build an RDD are naturally occurring elements that are naturally radioactive. Other components are solely available to radiological work sites such as reactor facilities, isotope production, research and universities, weapons production facilities, and other commercial and industrial facilities.

The lands surrounding ISIL are vast, and currently a large majority are under militant occupation.  The probability that materials of this nature can move about the region is very high.  Paid courier services, as well as stolen armed and armored transportation vehicles  may assist in this endeavor.

\subsubsection{???????? Identify possible and likely sources of the top 10 RDD materials identified in section 2.2 that ISIL might purchase or steal. Also identify possible sources of these materials within territory that ISIL controls (such as hospitals and universities). Discuss any current export controls on these materials, and how they might inhibit the acquisition of significant quantities of materials. Discuss the orphan source program, or any other international institutions which regulate the creation and distribution of radioactive sources. (To be added by 4/19; possible add map of sources?) ?????????????}

 \subsection{Flow of fission weapons/materials}

Merge with above subsection?

@@

\section{Weaponization and Delivery}

>>>>

Assuming that the White House institutionalizes this role as a Deputy National Security Advisor, the costs associated with this recommendation equal the salary and benefits of this person as well as a small staff. There will also be costs associated with the initiatives created by this person and other staff time that is devoted to those initiatives. Creating this role is an important step in crafting a strategic plan that integrates departmental efforts. The downside to this recommendation is that the efforts of this person are likely to be hampered by turf wars inherent in integrating departmental efforts; however, this difficulty can be mitigated by an approach that seeks to provide departments with a framework for evaluating and adjusting their efforts in the context of the broader effort, rather than mandating changes from the top down \cite{Levi2008}. This way, it promotes cross-agency relationships and personnel become better equipped at sharing and comparing methods of collecting and analyzing information.  Agencies should be less structured on rules and regulations and instead focus on how to best adapt to changing circumstances. A combination of minimal oversight and conflicting missions due the independent nature of these agencies \cite{Zegart2005} require that this person implement a uniformity of standards to decrease costs and simplify the flow of information.

@@

% As ISIL carves its traction in Iraq and Syria, the two countries become an immediate concern for help by the U.S.. Strengthening Iraq and Syria could be among the first line of defense against ISIL. As U.S. has an improving bilateral relations with Iraq, Iraq become an easier partner for U.S. to work with in reigning ISIL. The Strategic Framework Agreement for a Relationship of Friendship and Cooperation between the U.S. and the Republic of Iraq guides the overall political, economic, cultural, and security ties with Iraq.  This agreement is designed to help the Iraqi people stand on their own and reinforce Iraqi sovereignty, while protecting U.S. interests in the Middle East. 
% 
% However, the country's relation with Syria is not black-and-white as we would like it to be. The U.S.-Syria relationship is negative after having suspended diplomatic relations in 2014. The U.S. has also imposed economic sanction to Syria since 2004 due to the country's involvement in harboring terrorist groups and the meddling of Lebanon's political affairs including the assassination of Lebanese Prime Minister Rafik Hariri. The U.S. deeply opposed to the ongoing Syrian Civil war and President Bashar al-Assad's regime, specifically regarding the country's use and stockpile of chemical and biological weapons. The U.S. has floated the idea to train the Syrian military opposition, and even considering to train the Syrian military itself in order to fight ISIL, nevertheless, President Assad viewed the proposal as infringing on their country's sovereignty. Despite the two countries differences, they actually both share a common enemy, and it is hope this fact can be harnessed to U.S.' advantage.

@@

% \subsubsection{RDD}
%  Some of the materials required to build an RDD are naturally occurring elements that are naturally radioactive. Other components are solely available to radiological work sites such as reactor facilities, isotope production, research and universities, weapons production facilities, and other commercial and industrial facilities. 
% 
% The lands surrounding ISIL are vast, and currently a large majority are under militant occupation.  The probability that materials of this nature can move about the region is very high.  Paid courier services, as well as stolen armed and armored transportation vehicles  may assist in this endeavor.
% 
% \subsubsection{???????? Identify possible and likely sources of the top 10 RDD materials identified in section 2.2 that ISIL might purchase or steal. Also identify possible sources of these materials within territory that ISIL controls (such as hospitals and universities). Discuss any current export controls on these materials, and how they might inhibit the acquisition of significant quantities of materials. Discuss the orphan source program, or any other international institutions which regulate the creation and distribution of radioactive sources. (To be added by 4/19; possible add map of sources?) ?????????????}

@@

@@

\section{Weaponization \& Delivery}
